---
title: "Why dataviz in R?"
output: html_notebook
bibliography: references.bib
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
library(palmerpenguins)
library(countrycode)
library(ggrepel)
library(plotly)
```

Data visualization is a key part of data science. We can use dataviz in two main ways:

1.  **During the research process** Quick sketches, that highlight the internal patterns of data, and help us understand the phenomena.
2.  **Reporting results** Detailed visualization that help us in the storytelling of our publication.

Also, there are two main type of visualizations:

1.  **Static** For papers, posters, and printed stuff
2.  **Interactive** For computer-supported materials

So, we need a dataviz language that allows:

1.  Easy sketching
2.  Detail tailoring as needed
3.  Good static publication-level visualizations
4.  Good interactive visualizations.

We could also add:

5.  Easy to understand
6.  Reusable code
7.  With a large gamma of options
8.  With pre-made thematic plots

- *In my personal experience*, the combination ggplot [@ggplot] + plotly [@plotly] + shiny [@shiny] covers all these conditions
- This that not mean the it is the _only_ or _best_ option for any of the above points, but the one that better handles the trade-off between all.
- For example, the best known tool for interactive visualizations is neither R nor python, but the [D3 library](https://d3js.org/) for javascript.


# GGPLOT

In `ggplot` we think of plots as a succession of layers, which are built one at a time.    

- The __```+```__  operator allows us to add new layers to the plot.

- The ```ggplot()``` command allows us to define the __data source__ and the __variables__ that will determine the axes of the plot (x,y), as well as the color and shape of the lines or points, etc. All the mapped attributes go inside the `aes()` 

- The successive layers allow us to define:
   
    - One or more types of graphics (geometries):
        - ```geom_col()```, 
        - ```geom_line()```
        - ```geom_point()```
        - ```geom_boxplot()```
    - titles and axis names ```labs()```
    - plot styling ```theme()```
    - axis scalses ```scale_y_continuous```,```scale_x_discrete``` 
    - facetting ```facet_wrap()```,```facet_grid()```


for example 

We can make a quick sketch in two lines of code

```{r message=FALSE, warning=FALSE}
ggplot(penguins, aes(x = flipper_length_mm,y = body_mass_g,color = sex)) +
  geom_point() 
```
But we can also add as much detail as we want

```{r message=FALSE, warning=FALSE}
plt <- penguins %>% 
  filter(!is.na(sex)) %>% 
  ggplot(., aes(x = flipper_length_mm,
                       y = body_mass_g,
                       color = sex)) +
  geom_point() +
  facet_wrap(~species) +
  theme_minimal() +
  scale_color_manual(values = c("darkorange","cyan4")) +
  labs(title = "Penguin flipper and body mass",
       subtitle = "Dimensions for male and female Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin sex",
       shape = 'Island') +
  theme(legend.position = "bottom",
        legend.background = element_rect(fill = "white", color = NA),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

plt
```

- The pre-defined themes, `theme_*` allows to quicky adjust the style of the plot, while with `theme` let us correct every possible detail. 

## interactivity

-Making this an interactive plot is only one extra line of code


```{r, fig.width=10, fig.height=5}
ggplotly(plt)
```



##  [ggplot extensions](https://exts.ggplot2.tidyverse.org/gallery/).

The ggplot library in turn has many other libraries that extend its potential. Among my favorites are:


- [gganimate](https://gganimate.com/): for animated plots.
- [ggridge](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html): for faceted density plots
- [ggally](https://ggobi.github.io/ggally/): for grids of plots and specific visualizations
- [treemapify](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) for treemaps


for example

```{r message=FALSE, warning=FALSE}
library(GGally)

penguins %>% 
  select(species:bill_depth_mm) %>% 
ggpairs(mapping = aes(color = species))
```


```{r}
library(ggridges)

ggplot(penguins, aes(x = bill_length_mm, y = species, fill=species)) + 
  geom_density_ridges()
```


```{r}
library(corrr)

penguins %>% 
  select(bill_length_mm:body_mass_g) %>% 
correlate(.) %>%
  network_plot(.)
```


- This allows for very specific plot types to be a one-line thing, without loosing the ability of adding our own details later (is not a black-box)




## Shiny

- Shiny is a package that allows to make interactive dashboards, where the user defines the parameters of the plots. 

- For me, this is a great tool in two workflows:

1. As Supporting materials to a paper, to increase the engagement of the readers: i.e. https://ldaglobaltrade.uni.lu/dashboard/
2. During the project development, to communicate intermediate results with the team: i.e. https://sciencebias.uni.lu/dev/rg_app/ (user:'tmp', passwd: 'user', please do not disclose)

---------